Optimal (Multiway) Spatial Joins

### 2024-06-01 msj.tex
yufei

Now, suppose that each interval of $S$ is associated with a real-valued {\em weight}. Given a real value $q$, a {\em stabbing max} query returns the maximum weight of all the intervals in $S$ covering $q$ (if no such intervals exist, the query returns $-\infty$). We can store $S$ in a structure of \cite{aak+12} using $O(|S|)$ space that can answer such a query in $O(\log |S|)$ time. The structure supports an insertion or deletion on $S$ in $O(\log |S|)$ amortized time.

@@

\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{B}}$.} Sort the set of y-coordinates defined by all the segments of $H$ and $V$. Next, sweep (conceptually) a horizontal line $\ell$ from $y = -\infty$ to $y = \infty$. At all times, maintain the set $V_\ell$ of segments in $V$ that intersect with $\ell$. Let $S_\ell$ be the set of x-coordinates of the segments in $V_\ell$; we store $S_\ell$ in a binary search tree (BST) $\T$. Specifically, when $\ell$ hits the lower  (resp., upper) endpoint of a vertical segment $v = x \times [y_1, y_2]$ of $V$, we insert (resp., delete) the value $x$ into (resp., from) $\T$, which can be done in $O(\log n)$ amortized time. When $\ell$ hits a horizontal segment $h = [x_1, x_2] \times y$ of $H$, search $\T$ to determine if the successor $x'$ of $x_1$ in $S_\ell$. If $x' \le x_2$, then we output a pair $(h, p)$, where $p$ is the point $(x', y)$. The overall running time is $O(n \log n)$.

@@

When $\ell$ hits a horizontal segment $h = [z_1, z_2] \times y$ of $P$, perform a containment query on $\T$ to find all the intervals in $S_\ell$ that are contained by $[z_1, z_2]$; if $K_h$ is the number of such intervals, this retrieval takes $O(\log n + K_h)$ time. For each retrieved interval $[x_1, x_2]$, we also obtain its host rectangle $r$ (stored along with $[x_1, x_2]$ in $\T$). Observe that $h$ is the lowest segment in $H$ that crosses $r$; we therefore output $(r, h)$. After that, $r$ is marked as inactive, and accordingly, its x-range $[x_1, x_2]$ is deleted from $T$ in $O(\log n)$ time. As $r$ will remain inactive in the rest of the execution, its x-range will not be retrieved again by another containment query in the future.

>>>>

Now, let us assume that each interval of $S$ is associated with a real-valued {\em weight}. Given a real value $q$, a {\em stabbing max} query returns the maximum weight of all the intervals in $S$ covering $q$ (if no such intervals exist, the query returns $-\infty$). We can store $S$ in a structure of \cite{aak+12} using $O(|S|)$ space that can answer such a query in $O(\log |S|)$ time. The structure supports an insertion or deletion on $S$ in $O(\log |S|)$ amortized time.

@@

\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{B}}$.} Sort the set of y-coordinates defined by all the segments of $H$ and $V$. Next, sweep (conceptually) a horizontal line $\ell$ from $y = -\infty$ to $y = \infty$. At all times, maintain the set $V_\ell$ of segments in $V$ that intersect with $\ell$. Let $S_\ell$ be the set of x-coordinates of the segments in $V_\ell$; we store $S_\ell$ in a binary search tree (BST) $\T$. Specifically, when $\ell$ hits the lower  (resp., upper) endpoint of a vertical segment $v = x \times [y_1, y_2]$ of $V$, we insert (resp., delete) the value $x$ into (resp., from) $\T$, which can be done in $O(\log n)$ amortized time. When $\ell$ hits a horizontal segment $h = [x_1, x_2] \times y$ of $H$, search $\T$ to determine if the successor $x'$ of $x_1$ in $S_\ell$. If $x' \le x_2$, then we output a pair $(h, p)$, where $p$ is the point $(x', y)$; otherwise, output nothing for $h$. The overall running time is $O(n \log n)$.

@@

When $\ell$ hits a horizontal segment $h = [z_1, z_2] \times y$ of $P$, perform a containment query on $\T$ to find all the intervals in $S_\ell$ that are contained by $[z_1, z_2]$; if $K_h$ is the number of such intervals, this retrieval takes $O(\log n + K_h)$ time. For each retrieved interval $[x_1, x_2]$, we also obtain its host rectangle $r$ (stored along with $[x_1, x_2]$ in $\T$). As can be verified shortly, $h$ is the lowest segment in $H$ that crosses $r$; we therefore output $(r, h)$. After that, $r$ is marked as inactive, and accordingly, its x-range $[x_1, x_2]$ is deleted from $T$ in $O(\log n)$ time. As $r$ will remain inactive in the rest of the execution, its x-range will not be retrieved again by another containment query in the future. This implies that $h$ is indeed the lowest segment in $H$ crossing $r$.
